---
title: "Weekly Mosquito Surveillance and Forecast Report"
author: Florida Department of Health in Alachua County
output:
  pdf_document:
    highlight: zenburn
    toc: yes
  html_document:
    theme: united
    toc: yes
graphics: yes
header-includes:
     - \usepackage{graphicx}
     - \usepackage{grffile}
---




```{r, echo = FALSE}
library(knitr)
opts_chunk$set(comment=NA, warning = FALSE, message=FALSE, echo = FALSE)
```


```{r}
# Ordinal function:
ordinal <- function(number){
  number <- as.numeric(number)
  suffix <- if (number %% 100 %~% paste(11:13, collapse = "|")) {
    'th'
  } else {
    switch(number %% 10 + 1, 'th', 'st', 'nd', 'rd', 'th')             
  }
  sprintf("%d%s", number, suffix)
}

`%~%` <- function (x, rx){
    base:::regexpr(rx, x) > 0 
}
```

```{r, include = FALSE}
# wd <- '/home/joebrew/Documents/mosquito15/'
# setwd(wd)
library(plyr)
library(dplyr)
library(png)
library(grid)
library(ggplot2)
library(rgdal)
library(ggmap)
library(tidyr)
library(randomForest)
library(RColorBrewer)
load('../checkpoint.RData')

# create a blank ggplot theme
theme_opts <- list(theme(panel.grid.minor = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.background = element_blank(),
                        plot.background = element_rect(fill="white"),
                        panel.border = element_blank(),
                        axis.line = element_blank(),
                        axis.text.x = element_blank(),
                        axis.text.y = element_blank(),
                        axis.ticks = element_blank(),
                        axis.title.x = element_blank(),
                        axis.title.y = element_blank(),
                        plot.title = element_text(size=22)))

# Read in Alachua boundary
boundary <- readOGR('../geo_data/alachua_boundary/', 'Alachua_Boundary')

# Trap locations
trap_locations <- read.csv('../geo_data/traplatlong.csv')


```

```{r, fig = TRUE, echo=FALSE, fig.align = 'center', warning = FALSE, message = FALSE}
# Get most recent trapping numbers
left <- trap_locations
right <- df[which(df$date == recent),c('date', 'trap', 'n')]
right$trap <- as.numeric(as.character(right$trap))
temp <- left_join(x = left, y = right)
temp$mosquitoes <- temp$n

ggplot(boundary, aes(long, lat)) +
  geom_polygon(color = 'darkgrey', fill = 'grey', alpha = 0.3) + 
  coord_equal() +
  theme_opts +
  geom_point(data = temp, aes(x = long, y = lat, size = mosquitoes), pch = 16,
             alpha = 0.5, color = 'black') +
  scale_size(range = c(2,20)) +
  theme(legend.position="none") 
# plot(boundary)
```

\begin{center}
Joe Brew \\
Ben Brew \\
Yoni Teitelbaum
\end{center}

```{r, fig = TRUE, fig.height = 0.5, echo=FALSE, fig.align = 'center'}
img <- readPNG('../images/logo.png')
grid.raster(img)
```

\newpage

# Overview

```{r}
# Get percentile
percentiles <- quantile(ts$n / ts$n_traps, probs = seq(0, 1, 0.01))
difs <- abs((ts$n / ts$n_traps)[nrow(ts)] - percentiles)
num <- as.numeric(gsub('%', '', names(which.min(difs))))

```

The most recent trap collection was on `r format(recent, '%B %d, %Y')`. The `r ts$n[which(ts$date == recent)]` mosquitoes trapped (approximately `r round(ts$n[which(ts$date == recent)] / ts$n_traps[which(ts$date == recent)])` per trap) is at the `r num` percentile of all historical trappings.

```{r, fig.height = 3, fig.width = 5, fig.align = 'center'}
temp <- ts
temp$p <- temp$n / temp$n_traps
val <- temp$p[nrow(temp)]
g <- ggplot(data = temp, aes(p))
g + 
  geom_density(fill = 'blue', alpha = 0.5) +
  scale_x_log10() +
  xlab('Mosquitoes per trap (weekly) (log scale)') +
  theme_bw() +
  geom_vline(xintercept = val, lwd = 2, alpha = 0.5)
```

As predicted, the late September spike in mosquitoes has subsided  Our model predicts continued medium-low levels throughout the rest of October.

```{r}
past_temp <- ts[which(ts$date > (Sys.Date() - 60)),]
future_temp <- future_agg

future_temp <- rbind.fill(past_temp[nrow(past_temp),], future_temp)

future_temp <- future_temp[which(future_temp$n > 0),]

g <- ggplot(data = future_temp) + 
  aes(x = date, y = n) +
  theme_bw() +
#   geom_line(aes(x = date, y = lower ), color = 'red') +
#   geom_line(aes(x = date, y = upper ), color = 'red') +
  geom_ribbon(aes(x = date, y = upper, ymin = lower, ymax = upper), alpha = 0.1, fill = 'darkred') +
  geom_line(color = 'darkred', lwd = 1, lty = 2, alpha = 0.5) + 
  xlab('Date') + 
  ylab('Mosquitoes') +
  geom_line(data = past_temp, aes(date, n), 
            color = 'darkred', lwd = 2) +
  annotate('text', 
           x= c(min(past_temp$date) + 5, max(future_temp$date) -5), 
           y = max(past_temp$n)*1.2, 
           label =  c('Observed', 'Predicted'), 
           size = 5,
           alpha = 0.5)
g 
```

# Historical perspective

Numbers so far this summer have been relatively low compared to previous summers. The recent spike is similar in magnitude to the 2014 spikes.   

```{r, fig.height = 6}
temp <- ts
temp$p <- temp$n / temp$n_traps
temp$day <- as.numeric(format(temp$date, '%j'))
temp$year <- as.numeric(format(temp$date, '%Y'))
temp <- temp[,c('year', 'day', 'p')]

g <- ggplot(data = temp, aes(x = day, y = p))
g + geom_point() +
  geom_line() +
  theme_bw() + 
  facet_grid(year ~. ) +
  xlab('Day number') +
  ylab('Mosquitoes per trap')
```


\newpage 

# Disease details  

Vectors of diseases of concern appear relatively evenly distributed throughout the county. In recent weeks, the number of mosquitoes capable of carrying SLEV increased substantially but no longer remains as elevated as during September.

```{r, fig.height = 8, fig.width = 8}
# Get disease types
temp <- left_join(x = raw_data, y = diseases)

# # Get only recent trapping
# temp <- temp[which(temp$date == recent),]

# Get last six trapping

trappings <- unique(sort(temp$date))
trappings <- trappings[(length(trappings) - 5):length(trappings)]
temp <- temp[which(temp$date %in% trappings),]

# # Get by date, trap and disease
temp_grouped <- temp %>%
  group_by(date, trap) %>%
  summarise(wnv = sum(n[wnv], na.rm = TRUE),
            slev = sum(n[slev], na.rm = TRUE),
            eeev = sum(n[eeev], na.rm = TRUE),
            chik = sum(n[chik], na.rm = TRUE),
            malaria = sum(n[malaria], na.rm = TRUE),
            dengue = sum(n[dengue], na.rm = TRUE),
            yellow = sum(n[yellow], na.rm = TRUE))

#Gather from wide to long
temp_long <- gather(data = temp_grouped, 
                    key = disease,
                    value = n,
                    wnv:yellow)

# Join to trap_locations
temp_long <- left_join(x = temp_long, 
                  y = trap_locations[,which(!names(trap_locations) %in% c('name'))], 
                  by = 'trap')

cols <- brewer.pal(7, 'Dark2')

g <- ggplot(boundary, aes(long, lat)) +
  geom_polygon(color = 'darkgrey', fill = 'grey', alpha = 0.3) + 
  coord_equal() +
  theme_opts +
  geom_point(data = temp_long, aes(x = long, y = lat, size = n, color = disease), pch = 16,
             alpha = 0.8) +
  scale_size(range = c(2,10)) +
  theme(legend.position="none") +
  facet_grid(disease ~ date, scales = 'fixed') +
  scale_colour_manual(values = cols)
g
```

\newpage

# Forecast 

Over the next two weeks, we predict the mosquito population to grow slightly, with the greatest growth ocurring in the east of the county.

```{r, fig.width = 8, fig.height = 4}
left <- trap_locations
right <- future[which(future$date > recent), c('date', 'trap', 'predicted')]
right$trap <- as.numeric(as.character(right$trap))
temp <- left_join(x = left, y = right)
temp$mosquitoes <- temp$predicted
temp <- temp[which(temp$date %in% as.Date(c('2015-10-30', '2015-11-02', '2015-11-05'))),] ##############!!!!!!!!!!!!!!!!!!

map <- get_map(location = c(lon = mean(temp$long), 
                            lat = mean(temp$lat)),
               source = 'stamen',
               maptype = 'toner',
               crop = FALSE,
               color = 'bw',
               zoom = 11)
ggmap(map) +
  geom_point(data = temp, aes(x = long, y = lat, size = mosquitoes), pch = 16,
             color = 'red', alpha = 0.5) +
  #scale_size_area(range = c(2,20)) +
  ggtitle('Predicted') +
  facet_grid(~ date, scales = 'fixed') 
```


```{r, fig.align = 'center', fig.width = 6, fig.height = 4, eval = TRUE}
left <- trap_locations
right <- future[which(future$date > recent), c('date', 'trap', 'predicted')]
right$trap <- as.numeric(as.character(right$trap))
temp <- left_join(x = left, y = right)
temp$mosquitoes <- temp$predicted
temp <- temp[which(temp$date >= recent),]
temp <- temp %>%
  group_by(trap, lat, long, name) %>%
  summarise(predicted = mean(predicted, na.rm = TRUE))

map <- get_map(location = c(lon = mean(temp$long), 
                            lat = mean(temp$lat)),
               source = 'google',
               maptype = 'terrain',
               crop = FALSE,
               color = 'bw',
               zoom = 10)
ggmap(map) +
  geom_point(data = temp, aes(x = long, y = lat, size = predicted), pch = 16,
             color = 'red', alpha = 0.5) +
  #scale_size_area(range = c(2,20)) +
  ggtitle('Predicted') +
  xlab('Longitude') + 
  ylab('Latitude')
```

\newpage

# Model details

The forecast model has undergone several improvements since last year.  The inputs have been substantially expanded (now taking into account humidity, precipitation, wind speed, temperature and its fluctuations), and the model has been improved (in lieu of linear regression, the predictive model now employs ensemble machine learning methods).  

The model is cross-validated and backtesting confirms a decent fit.

```{r}
temp <- df
temp <- temp[complete.cases(temp),]
temp$predicted <- predict(fit)
temp <- temp[,c('date','predicted', 'n', 'trap')]
temp <- temp %>%
  group_by(date) %>%
  summarise(predicted = sum(predicted, na.rm = TRUE),
            n = sum(n, na.rm = TRUE))

g <- ggplot(data = temp, aes(x = predicted, y = n))
g + geom_point(color = 'blue', alpha = 0.5) +
  labs(x = 'Mosquitoes predicted', y = 'Actual mosquitoes', title = 'Model performance') +
  theme_bw() +
  geom_smooth()
```
